Forcasting US Elections using Polls-of-Polls

### 2024-10-22 paper/paper.qmd
fixed typo and add link to survey
downloaded some other packages.
finished summary statistics and data section + added all of the LLM usage instances and documentation

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

@@

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

@@

# Emerson College Polling Methodology (October 14-16, 2024)

## Overview
Emerson College Polling, known for its extensive online polling methods, conducted a survey from October 14 to 16, 2024, targeting 1,000 likely voters. The poll measured voter preferences between Kamala Harris and Donald Trump, revealing a near tie with Harris at 50% and Trump at 49%.

## Population, Frame, and Sample
The target population is likely voters in the U.S. elections. These voters are determined by the voters' likeliness to vote in the upcoming elections and the voter's voting history. Both of these factors are self-reported.

The sampling frame is the voters that are on the Aristotle list and on the online panel provided by CINT. Aristotle is an online supplier of voter files and CINT is a software company that connects researchers, such as Emerson, to online panels. This approach helps ensure a representative sample by accurately reflecting the electorate.

The sample size is the 1,000 likely voters chosen arbitrarily from the sampling frame. Likely voters are determined by a combination of voter history, registration status, and demographic data. These factors are self-reported.

## Sample Recruitment and Approach
Emerson College used a mixed-mode sampling approach for its polling. Specifically, three primary modes were used to collect data for its October 14-16, 2024 poll:

- MMS-to-web text surveys: Respondents were contacted via text messages sent to cell phones. These messages directed participants to complete the survey online. This method was conducted using voter lists provided by Aristotle.

- Interactive Voice Response (IVR): Landlines were targeted using automated phone calls where respondents could answer questions using their keypads. The contact information for this was also sourced from Aristotle’s voter lists.

- Online Panel from CINT: Emerson utilized a pre-screened, opt-in online panel of voters provided by CINT.

## Advantages and Trade-offs in Mixed-Mode Sampling
The multi-mode sampling approach used in this poll has several advantages:

- Wider demographic reach: Having a combination of data collection methods allows Emerson to gather responses from a broader demographic group. In this polling, MMS-to-web surveys mostly gather responses from younger voters and those who primarily use mobile devices. IVR target older and rural voters, who are more likely to respond via landlines. Online panels is more efficient in reaching tech-savvy individuals.

- Cost-efficient: Automated technologies reduce the cost compared to live interviews. In a polling context, this allows for a larger sample size.

However, the trade-offs include:

- Coverage bias: IVR only reaches individuals with landlines, which may skew results towards older demographics.

- Self-selection bias: Respondents in online panels tend to be more engaged, which can lead to over-representation of politically active individuals.

## Non-response Handling
Emerson addresses non-response using a weighting system. The data is adjusted based on key demographic variables like age, gender, race, education, and party affiliation, making the sample more representative of the electorate. Weights are applied to compensate for under- or over-represented groups, ensuring balanced results.

## Questionnaire Design
The Emerson survey features a straightforward, efficient design focused on key issues and voter preferences. Strengths of the questionnaire include:

- Simplicity: Clear, concise questions makes the survey easy for respondents to understand and to provide accurate answers.

- Topical relevance: The poll zeroes in on vote preference and demographic splits, delivering timely insights during election cycles.

However, weaknesses include:

- Lack of nuance: While the survey captures basic preferences, it does not delve deeply into voters' motivations or the factors driving their choices.

- Mode limitations: Different survey modes can affect how respondents engage with the questions. This can potentially influence the depth and accuracy of voter responses.

## Conclusion
Emerson College Polling's mixed-mode methodology is effective in balancing cost with broad demographic reach. Its weighting techniques contribute to reliable and accurate results despite non-responses. While the use of multiple survey modes increases population representation, possible biases—such as lack of nuance and mode effects need to be taken into consideration when interpreting results.

In sum, this method is a strong predictor of election outcomes and continues to adapt to the evolving challenges of modern polling.

# Idealized Methodology

## Overview
This methodology outlines an election forecasting plan with a budget of $100,000. Heavily inspired by TIPP Insights and Emerson College's hybrid approaches, the survey methodology balances cost-efficiency and data accuracy by using different modes of data collection and wide demographic reach.

This forecasting plan will predict the popular vote and Electoral College outcomes, emphasizing key swing states and underrepresented voter groups.

## Sampling Approach
This idealized methodology uses a multi-mode hybrid sampling approach. This allows the use of both probability-based and non-probability-based methods:

- Probability-based methods:

  - Live phone interviews (landline and cell): As used by TIPP, live phone interviews is a direct and personal method of collecting responses. Voter lists would be sourced from Aristotle to ensure we contact likely voters. This mode reaches older voters and voters with less internet connection.

  - Interactive Voice Response (IVR): Similar to Emerson’s approach, IVR for landline users balances costs and ensures demographic representation, particularly among older or less tech-savvy populations. This method captures quick responses and is less expensive than live interviews.

- Non-probability methods:

  - MMS-to-web (mobile): Following Emerson’s MMS-to-web strategy, we would send multimedia messages to likely voters’ cell phones, encouraging them to complete an online survey. This method works well for younger, mobile-first voters, helping fill in gaps where traditional methods might miss.

  - Online panel recruitment via CINT: To mirror TIPP’s online panel approach, a portion of the survey would be conducted via an opt-in panel of pre-screened voters from CINT, capturing tech-savvy respondents. CINT’s platform ensures that different voter groups, including hard-to-reach groups, are represented in the polling.

## Recruitment Strategy
By using Aristotle’s voter lists and CINT's platform as a sampling frame, the sample consists of likely voters based on registration status and past voting behavior. Quota sampling would be used to match the demographic profile of the U.S. electorate. This ensures that respondents in the sample reflect key demographic variables such as age, gender, race, and education level.

We would oversample key swing states like Pennsylvania, Wisconsin, and Florida. Key swing states are states where there is no overwhelming support for a particular candidate or political party. The outcome in these states can often determine the overall result of the election, making them critical in securing a win in the Electoral College. Oversampling these states allows a more accurate forecasting in these battlegrounds, which are critical for predicting the Electoral College outcome.

>>>>

---
title: Forcasting US Elections using Polls-of-Polls
subtitle: "My subtitle if needed"
author: 
  - Shamayla Durrin
  - Denise Chang 
  - Krishna Kumar
thanks: "Code and data are available at: [https://github.com/krishnak30/US_elections](https://github.com/krishnak30/US_elections)."

date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Data

## Data Set Overview

add a summary statistics table here maybe?

```{r}
library(arrow)
analysis_data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))

#head(analysis_data, 20)
colnames(analysis_data)
#print(nrow(analysis_data))
```

## Variables

The data set contains a wide range of variables that are essential for understanding polling outcomes. The following are key variables:

-   **`poll_id`**: Unique identifier for each poll conducted.

-   **`pollster`**: The polling organization that conducted the poll (e.g., YouGov, TIPP).

-   **`numeric_grade`**: A numeric rating indicating the pollster’s quality or reliability (e.g., 1.8).

-   **`pollscore`**: A score representing the reliability of the pollster in terms of error and bias.

-   **`state`**: The state where the poll was conducted (or "Not Applicable" for national polls).

-   **`start_date` / `end_date`**: Dates during which the poll was conducted.

-   **`national`**: A binary variable which determines where a poll was national (1) or state (0)

-   **`sample_size`**: The number of respondents in the poll.

-   **`population`**: The group surveyed (e.g., likely voters or registered voters).

-   **`candidate_name`**: The name of the candidate being polled (e.g., Kamala Harris, Donald Trump).

-   **`pct`**: The percentage of support for the candidate in the poll.

-   **`methodology`**: The method used to conduct the poll (e.g. online panel)

-   **`num_support`**: The number of people supporting a candidate in a poll

## High Level Data Cleaning

blah blah blah

## Summary Statistics

### 1) Average Pollster Quality by Poll Type (pollscore vs state/national): mean, median, mode, min, max

```{r}
#| warning: false
#| message: false
#| echo: false

library(dplyr)

# Generate summary statistics for national and state polls
summary_stats <- analysis_data %>%
  filter(!is.na(pollscore)) %>%  # Exclude rows with missing pollscore
  group_by(national) %>%         # Group by the binary 'national' column (1 for national, 0 for state)
  summarize(
    Count = n(),
    Min = round(min(pollscore), 2),
    `1st Qu.` = round(quantile(pollscore, 0.25), 2),
    Median = round(median(pollscore), 2),
    Mean = round(mean(pollscore), 2),
    `3rd Qu.` = round(quantile(pollscore, 0.75), 2),
    Max = round(max(pollscore), 2),
    .groups = 'drop'  # Drop grouping structure for cleaner output
  ) %>%
  # Rename 'national' to 'Poll Type' and replace values
  mutate(`Poll Type` = ifelse(national == 1, "National", "State")) %>%
  select(`Poll Type`, Count, Min, `1st Qu.`, Median, Mean, `3rd Qu.`, Max)  # Reorder columns

# Print the summary statistics
print(summary_stats)

```

As shown above, the summary statistics table **\[CROSS-REFERENCE HERE\]** offers an overview of poll scores for national and state polls, with 427 and 696 entries, respectively. Both categories feature a minimum poll score of -1.5, which a high level of validity and reliability. The first quartile shows national polls at -0.9, while state polls are slightly lower at -1.1. Median scores are -0.3 for national polls and -0.4 for state polls, suggesting that half of the polls in both categories yield moderately negative scores, reinforcing a trend of favorable sentiment. The mean score is -0.39 for national polls, indicating a less reliable perception compared to the -0.54 mean for state polls. The third quartile for national and state polls (-0.3) indicates that the upper 25% of scores are slightly above neutral. The maximum score of 1.7 for national polls reflects a notable negative perception for at least one candidate, while the maximum for state polls is 0.6. Overall, these statistics illustrate a strong reliability of pollsters considering their errors and bias.

### 2) Distribution of Sample Sizes

```{r}
#| warning: false
#| message: false
#| echo: false

library(ggplot2)
library(dplyr)
library(gridExtra)

# Adjust the sample size range categories
analysis_data <- analysis_data %>%
  mutate(sample_size_range = case_when(
    sample_size <= 6000 ~ "0-6,000",
    sample_size > 6000 & sample_size <= 13000 ~ "6,001-13,000",
    TRUE ~ "Above 13,000"  # Catch any values above 13,000
  ))

# Plot for 0-6,000
p1 <- ggplot(analysis_data %>% filter(sample_size <= 6000), aes(x = sample_size)) +
  geom_histogram(binwidth = 500, fill = "lightgreen", color = NA, alpha = 0.7) +
  labs(title = "Sample Sizes (0-6,000)",
       x = "Sample Size",
       y = "Count") +
  scale_x_continuous(limits = c(0, 6000)) +  # Set x limits for the first plot
  scale_y_continuous(limits = c(0, 500)) +  # Adjust y limits based on data
  theme_minimal() +
  theme(plot.title = element_text(size = 10, hjust = 0.5),  # **Center the title**
        axis.title.y = element_text(margin = margin(r = 15)))  # **Increased gap for y-axis label**

# Plot for 6,001 - 13,000
p2 <- ggplot(analysis_data %>% filter(sample_size > 6000 & sample_size <= 13000), aes(x = sample_size)) +
  geom_histogram(binwidth = 500, fill = "indianred", color = NA, alpha = 0.7) +
  labs(title = "Sample Sizes (6,001-13,000)",
       x = "Sample Size",
       y = "Count") +
  scale_x_continuous(limits = c(6001, 13000)) +  # Set x limits for the second plot
  scale_y_continuous(limits = c(0, 20)) +  # Adjust y limits based on data
  theme_minimal() +
  theme(plot.title = element_text(size = 10, hjust = 0.5),  # **Center the title**
        axis.title.y = element_text(margin = margin(r = 10)))  # **Increased gap for y-axis label**

# Arrange the two plots side by side
grid.arrange(p1, p2, ncol = 2)  # ncol=2 arranges the plots in one row

```

The histograms **\[CROSS-REFERENCE HERE\]** reveal significant trends in sample sizes across the different ranges, indicating a general tendency towards smaller samples in the polling data. **In Plot 1**, most of the values are concentrated on the left side, with very few values extending beyond the 4,500 mark. This clustering suggests that many polls are conducted with limited respondent numbers, which may compromise the reliability and representativeness of the results.

In **Plot 2**, the values are somewhat evenly divided between 6,000 and the maximum sample size of 12,084. This distribution shows a broader representation of sample sizes, indicating that while there are still smaller samples, there are also a notable number of polls that feature larger sample sizes. This contrast with Plot 1 highlights a shift towards more robust polling efforts.

The findings from these histograms suggest potential limitations and advantages in the accuracy and validity of the insights generated from the polling data. Smaller sample sizes are typically less desirable in polling, as they can yield less reliable estimates and increase the margin of error. Consequently, the predominance of smaller sample sizes in Plot 1 raises concerns about the robustness of the polling findings. However, the more even distribution in Plot 2 suggests that some polls conducted with larger sample sizes could enhance the reliability of the results.

Ultimately, the analysis emphasizes the need to consider sample sizes critically when interpreting polling data, as larger sample sizes tend to provide a better representation of the population's views. Thus, while many polls may fall short of ideal sample sizes, the existence of larger samples in the dataset can offer valuable insights into public opinion. **However, we need to determine an ideal and acceptable sample size that would be enough for the polling results to be considered reliable, before we conclude smaller samples as being unreliable.**

### 3) Summary Statistics Table for Numeric Variables

```{r}
#| warning: false
#| message: false
#| echo: false

library(dplyr)
library(modelsummary)
library(tools)  # Make sure to load the tools package for toTitleCase()

# Set options to use kableExtra for table formatting
options(modelsummary_factory_default = 'kableExtra')

# Create the summary statistics table for the specified numerical variables
summary_table <- analysis_data %>%
  select(numeric_grade, pollscore, sample_size, num_support) %>%
  # Remove underscores and capitalize the first letter of each word
  rename_with(~ toTitleCase(gsub("_", " ", .))) %>%
  datasummary_skim(
    fun_numeric = list(
      Mean = mean,
      SD = sd,
      Min. = min,
      Median = median,
      Max. = max
    )
  )

# Display the summary statistics table
print(summary_table)

```

The summary statistics **\[CROSS REFERENCE HERE\]** for the numeric variables provide significant insights into the polling data. The **numeric grade** has a mean of 2.3 and a standard deviation of 0.6, suggesting that most polls are graded around a middle value, though there is moderate variability. The range, with a minimum of 1.1 and a maximum of 3.0, indicates the presence of both low-quality and higher-quality polls within the dataset.

The **pollscore** averages at −0.5, with a standard deviation of 0.6. Since lower poll scores indicate more favorable outcomes for certain candidates or parties, this negative mean suggests that the polls are, on average, skewing positively towards their targets. The minimum of −1.5 and maximum of 1.7 further illustrate the wide distribution of poll results, highlighting the variability in outcomes across different polls.

Looking at the **sample size**, the mean of 1962.5, accompanied by a substantial standard deviation of 2608.0, reflects significant variability in respondent numbers. With a minimum sample size of 382 and a maximum of 12,084, it is evident that while many polls are conducted with relatively small sample sizes, a few achieve much larger respondent pools. This variability can enhance the reliability of findings, particularly for those polls with larger sample sizes.

Finally, the **number of supporters** reported has a mean of 949.2 and a standard deviation of 1290.0, indicating that while the average number of supporters is relatively high, there is considerable variability. The minimum value of 172 and maximum of 6042 underscores the diverse contexts in which these polls were conducted.

Overall, these summary statistics emphasize the variability and range present in the polling data, which may impact the reliability and interpretation of the findings. The mixture of low and high values in numeric grade, pollscore, and sample size points to the importance of careful analysis when drawing conclusions from these polling efforts.

### 4) Top 10 Pollsters and their Average Pollscores and Numeric Grades

```{r}
library(dplyr)

# Get the top 10 pollsters by frequency along with their average pollscore and numeric grade
top_pollsters <- analysis_data %>%
  group_by(pollster) %>%           # Group by the 'pollster' variable
  summarize(
    Count = n(),                   # Count the number of occurrences
    Average_Pollscore = round(mean(pollscore, na.rm = TRUE), 2),  # Average pollscore
    Average_Numeric_Grade = round(mean(numeric_grade, na.rm = TRUE), 2),  # Average numeric grade
    .groups = 'drop'               # Drop grouping structure for cleaner output
  ) %>%
  arrange(desc(Count)) %>%         # Arrange in descending order of count
  slice_head(n = 10)                # Select the top 10 pollsters

# Rename the columns to remove underscores and capitalize 'Pollster'
colnames(top_pollsters) <- c("Pollster", "Count", "Average Pollscore", "Average Numeric Grade")

# Print the top pollsters table
print(top_pollsters)

```

The examination of the top 10 pollsters in **\[CROSS REFERENCE HERE\]** highlights their significance in electoral forecasting. **Morning Consult**, with the highest volume of polls (235), shows a slightly optimistic average pollscore of −0.3, indicating some reliability and validity. Conversely, **Siena/NYT** stands out with the lowest average pollscore of −1.5 and the highest numeric grade of 3.0, suggesting a rigorous methodology that enhances its credibility in predicting electoral outcomes.

In contrast, **Redfield & Wilton Strategies** displays an average pollscore of 0.4, which may create a misleadingly pessimistic view of voter sentiment. This underscores the necessity of considering multiple polling sources for a nuanced understanding of electoral dynamics.

The presence of several pollsters with negative scores indicates a general trend towards specific candidates, which is crucial for shaping campaign strategies. Overall, these findings reinforce the critical role of trustworthy polling in influencing public perception and political strategy during elections.

### 5) Histogram of End Dates

## Graphs for Imp. Variables (relationships)

@@

Public support for political candidates is an essential part of understanding electoral dynamics for a variety of reasons. Public polls provide insights into the preferences and beliefs of certain groups of people, usually voters in specific states or districts, and sometimes to also correlate preferences to demographic information (race, age, neighborhood, etc).

The results help inform policy decisions, help track the likely outcomes of elections, help government officials and candidates decide where to invest time, funds, and resources to get the desired results. Polling organizations aim to capture data by surveying a representative sample of voters, free from government intervention.

Polling organizations conduct these surveys through various methods. For our particular data set, the following methods have been used to gauge public sentiment:

-   App panel

-   Online panel

-   Email

-   Online ad

-   IVR (interactive voice response)

-   Live phone

-   Text

-   Text-to-web

-   Probability panel

-   Mail-to-phone

Each pollster chooses a method(s) based on what they find to be most convenient, efficient, and providing the best results. Moreover, there are some variables which affect the validity and reliability of the polling results, such as sample size, methodology, population type, etc. Larger samples generally lead to more reliable estimates of public opinion. Some methodologies might be considered more reliable due to certain factors. The pollsters also use different sampling methods (random, stratified, etc) which can affect the validity of the polling results as well.

After collecting the required data, the polling organizations transform these results into more easily comprehensible variables such as numeric grade, poll score, transparency score, and a percentage. These scores are calculated through an extensive process of calculating excess error and bias, adjusting for race difficulty, calculating predictive error and bias, calculating pollster transparency, and combining everything into one single ‘POLLSCORE’, while accounting for luck as well.

Variables such as pollster rating and transparency score help validate the reliability of the pollster and the polling results they publish.

Each entry in the data set represents a glimpse of public opinion at a specific time and location, capturing measured support percentages alongside contextual variables such as pollster, methodology, and sample size. This extensive approach allows for an enhanced understanding of public support dynamics, facilitating informed analyses of electoral outcomes.

# Polls of Polls to Forecast Winning Candidate

## Forecasting Approach

The polls of polls methodology is widely used in election prediction as it aggregates multiple polls to provide a more reliable estimate of voter support, rather than relying on any single poll. The goal is to reduce errors and biases present in individual polls by using a weighted average of many different polls. In our approach, we calculate the weights based on several important variables. Sample size is included because larger polls are generally more reliable and have smaller margins of error. Pollscore (or pollster rating) reflects the empirical quality of a pollster, balancing both accuracy and bias, allowing us to give more weight to those with a proven track record. We also account for the pollscope, differentiating between national and state-level polls, as state polls may offer more localized insights into voter sentiment. We also use recency as more recent polls give us a more accurate estimate of voter sentiment.

However, we acknowledge that even with these adjustments, there may still be underlying biases from individual pollsters toward the candidates. To account for this, we will select a baseline pollster with a strong rating, high transparency, and minimal historical error, and regress the support percentage (pct) on pollster, including the same variables as controls whcih was used in the weighting method, such as sample size, pollscore and pollscope.Additionally, we will control for the state of the poll, as support levels can vary regionally. Therefore each pollster's coefficient in the regression model will represent the bias of that pollster relative to the baseline pollster. Controlling for these factors is crucial because it ensures that the differences between pollsters are not driven by other factors, such as differences in sample size or the types of polls they conduct, but by their inherent bias in support estimates. Once we have calculated the weights, we will use them to generate a weighted average of support for Kamala Harris and Donald Trump, ultimately comparing their weighted support to forecast the likely winner of the election.

## **Exploration of Control Variables**

### Investigating the Effect of Pollscore on Candidate Support

```{r}

# Define pollscore ranges (you can adjust the binwidth if needed)
cleaned_data <- analysis_data %>%
  mutate(pollscore_range = cut(pollscore, breaks = seq(-1.5, 1, by = 0.3), include.lowest = TRUE))

# Calculate the average support for each candidate within each pollscore range
average_support <- cleaned_data %>%
  group_by(pollscore_range, candidate_name) %>%
  summarise(avg_support = mean(pct, na.rm = TRUE)) %>%
  ungroup()

# Create a bar plot
ggplot(average_support, aes(x = pollscore_range, y = avg_support, fill = candidate_name)) +
  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate bars for each candidate
  scale_fill_manual(values = c("Donald Trump" = "indianred", "Kamala Harris" = "steelblue")) +  # Custom colors
  labs(title = "Average Support for Candidates by Pollscore Range",
       x = "Pollscore Range",
       y = "Average Support (%)",
       fill = "Candidate") +
  theme_minimal() +  # Minimal theme
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5)
  )

```

### Investigating the Effect of Poll Recency on Candidate Support

```{r}
#| warning: false
#| message: false
#| echo: false
# Custom colors for candidates
trump_color <- "indianred"
harris_color <- "steelblue"

# Create the plot with the requested adjustments (recency vs support)
ggplot(analysis_data, aes(x = recency, y = pct, color = candidate_name)) +
  geom_point(size = 0.5, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE) + # Smaller dots with slight transparency
  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
  theme_minimal() +  # Remove grey background
  theme(
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    legend.title = element_blank(),  # Remove "candidate_name" as title
    legend.position = "bottom",  # Move legend to the bottom
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.line = element_line(color = "black"),  # Add axis lines
    panel.border = element_blank(),  # Remove panel border
    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
    axis.line.x.top = element_blank()  # Remove top spine
  ) +
  labs(title = "Recency vs Support for Kamala Harris and Donald Trump",
       x = "Poll Recency (days since poll end date)", 
       y = "Support (%)") +
  facet_wrap(~ candidate_name, scales = "free_y")  # Separate plots for each candidate

```

### Investigate the effect of sample size on support estimates

```{r}
#| warning: false
#| message: false
#| echo: false

# Custom colors for candidates
trump_color <- "indianred"
harris_color <- "steelblue"

# Create the plot with the requested adjustments
ggplot(analysis_data, aes(x = sample_size, y = pct, color = candidate_name)) +
  geom_point(size = 0.5, alpha = 0.9) + # Smaller dots with slight transparency
  geom_smooth(method = "lm", se = FALSE)+

  scale_color_manual(values = c("Donald Trump" = trump_color, "Kamala Harris" = harris_color)) +  # Custom colors
  theme_minimal() +  # Remove grey background
  theme(
    panel.grid.major = element_blank(),  # Remove major gridlines
    panel.grid.minor = element_blank(),  # Remove minor gridlines
    legend.title = element_blank(),  # Remove "candidate_name" as title
    legend.position = "bottom",  # Move legend to the bottom
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.line = element_line(color = "black"),  # Add axis lines
    panel.border = element_blank(),  # Remove panel border
    axis.line.y.right = element_line(color = "black"),  # Add right spine for Kamala Harris
    axis.line.x.top = element_blank()  # Remove top spine
  ) +
  labs(title = "Sample Size vs Support for Kamala Harris and Donald Trump",
       x = "Sample Size", 
       y = "Support (%)") +
  facet_wrap(~ candidate_name, scales = "free_y")
```

### Distribution of support for Candidates by State

```{r}
# Calculate mean support for each candidate in each state
state_support <- analysis_data %>%
  group_by(state, candidate_name) %>%
  summarise(mean_support = mean(pct)) %>%
  ungroup()

# Pivot the data to have one row per state with separate columns for Harris and Trump
state_support_wide <- state_support %>%
  pivot_wider(names_from = candidate_name, values_from = mean_support)
```

```{r}
# Calculate mean support for each candidate in each state
state_support <- analysis_data %>%
  group_by(state, candidate_name) %>%
  summarise(mean_support = mean(pct))

# Calculate the proportion of support for Kamala Harris
state_support <- state_support %>%
  mutate(harris_proportion = Kamala_Harris / (Kamala_Harris + Donald_Trump))

```

@@

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```
